---
title: "Research"
author: "Anas"
date: "2024-07-01"
output: pdf_document
---

#Loading and processing data
```{r}
install.packages("ggplot2")
install.packages("dplyr")
library(ggplot2)
library(dplyr)
fileurl<- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
destfile<- "activity.zip"
download.file(fileurl,destfile, method= "curl")

unzip(zipfile="activity.zip")
activity<- read.csv("activity.csv")
activity$date<- as.Date(activity$date)
```

# claclulating the mean total number of steps taken per day
```{r}
totsteps<- tapply(activity$steps, activity$date, FUN = sum, na.rm= TRUE)
```

# Plotting the mean total number of steps taken per day
```{r}
qplot(totsteps, binwidith= 100, xlab= "total number of steps taken each day", ylab= "Frequency")
```

# claculating the mean and the median
```{r}
mean(totsteps, na.rm=TRUE)
median(totsteps, na.rm = TRUE)
```

# calculating the average daily activity
```{r}
steinter <- activity %>%
  group_by(interval) %>% 
  summarize(meansteps= mean(steps, na.rm=TRUE))
```

# ploting the average daily activity pattern
```{r}
ggplot(steinter, aes(x=interval, y= meansteps))+
  geom_line(size=1.5, color="red")+
  xlab("5 minute interval")+
  ylab("average number of steps taken")+
  theme_dark()
```
#5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
steinter[which.max(steinter$meansteps),]
```
# Imputing missing values
```{r}
misval<- is.na(activity)
sum(misval)
```

#Devise a strategy for filling in all of the missing values in the dataset
```{r}
activityNoNA <- activity  
for (i in 1:nrow(activity)){
        if(is.na(activity$steps[i])){
                activityNoNA$steps[i]<- steinter$meansteps[activityNoNA$interval[i] == steinter$interval]
        }
}

```
Make a histogram of the total number of steps take each day
```{r}
stepsPerDay <- activityNoNA %>%
        group_by(date) %>%
        summarize(sumsteps = sum(steps, na.rm = TRUE)) 

hist(stepsPerDay$sumsteps, main = "Histogram of Daily Steps", 
     col="lightblue", xlab="Steps")
```
#calculate the mean and median total number of steps taken per day
```{r}
meanPostNA <- round(mean(stepsPerDay$sumsteps), digits = 2)
medianPostNA <- round(median(stepsPerDay$sumsteps), digits = 2)
print(paste("The mean is: ", mean(meanPostNA)))
print(paste("The median is: ", median(meanPostNA)))
```

```{r}
NACompare <- data.frame(mean = c(mean(totsteps),meanPostNA),median = c(median(totsteps),medianPostNA))
rownames(NACompare) <- c("Pre NA Transformation", "Post NA Transformation")
print(NACompare)
```

#the difference in activity between weekdays and weekends
```{r}
activityDoW <- activityNoNA
activityDoW$date <- as.Date(activityDoW$date)
activityDoW$day <- ifelse(weekdays(activityDoW$date) %in% c("Saturday", "Sunday"), "weekend", "weekday")
activityDoW$day <- as.factor(activityDoW$day)
```

#panel plot
```{r}
activityWeekday <- filter(activityDoW, activityDoW$day == "weekday")
activityWeekend <- filter(activityDoW, activityDoW$day == "weekend")

activityWeekday <- activityWeekday %>%
        group_by(interval) %>%
        summarize(steps = mean(steps)) 
activityWeekday$day <- "weekday"

activityWeekend <- activityWeekend %>%
        group_by(interval) %>%
        summarize(steps = mean(steps)) 
activityWeekend$day <- "weekend"

wkdayWkend <- rbind(activityWeekday, activityWeekend)
wkdayWkend$day <- as.factor(wkdayWkend$day)


g <- ggplot (wkdayWkend, aes (interval, steps))
g + geom_line(size= 1.5, col="red") + facet_grid (day~.) + 
        theme(axis.text = element_text(size = 12),axis.title = element_text(size = 14)) + 
        labs(y = "Number of Steps") + labs(x = "Interval") + 
        ggtitle("Average Number of Steps: Weekday vs. Weekend") + 
        theme(plot.title = element_text(hjust = 0.5))
```
